---
title: "Assignment_questions"
author: "Lily Slater"
date: '2022-11-14'
output: html_document
---

### What states (of those I am willing to move to) have the highest paying data-related salaries? Differences between job sub-categoriES

#Loading datafile
```{r}
salary_data_states <- read.csv("~/Desktop/DS311/python_charmers/collab_files/salary_data_states.csv")
```


#Loading Packages 
```{r}
library(dplyr)
library(ggplot2) 
library(knitr)
```

#Exploring the datafile

```{r}
names(salary_data_states)
str(salary_data_states)
 
```



```{r}

#sub-setting columns of interest
data_interest = data.frame(salary_data_states[c("WORK_STATE", "JOB_TITLE_SUBGROUP", "PAID_WAGE_PER_YEAR")])
str(data_interest)
#excluding rows with na's
data_int = na.exclude(data_interest)
str(data_int)

#taking a look at jobs of interest [sub categories]
ff = data_int[data_int$JOB_TITLE_SUBGROUP == c("data analyst" ), ]
fg = data_int[data_int$JOB_TITLE_SUBGROUP == c( "data scientist"), ]

# group by work state
fm = distinct(ff, WORK_STATE, .keep_all = TRUE)
dfg = distinct(fg, WORK_STATE, .keep_all = TRUE)
#sorting
fm_sort <- fm[order(-fm$PAID_WAGE_PER_YEAR),]
da_df <- head(fm_sort)
dfg_sort <- dfg[order(-dfg$PAID_WAGE_PER_YEAR),]
da_dfg <- head(dfg_sort)

#Subsetting this data for states with high pay
da_dfcnn <- da_df[da_df$WORK_STATE %in% c("California", "New York", "New Jersey") ,]
dfg_dfcnn <- da_dfg[da_dfg$WORK_STATE %in% c("California", "New York", "New Jersey") ,]

```


Plotting this data to a scatterplot
```{r}

ggplot() +
  geom_point(data = da_dfcnn, aes(WORK_STATE, PAID_WAGE_PER_YEAR), color = "black", size = 1.5) +
  geom_point(data = dfg_dfcnn , aes(WORK_STATE, PAID_WAGE_PER_YEAR), color = "red" , size = 1.5) +
  ggtitle("Salary for Data Analysts (black) and Data Scientists (red) in Highest Paying States")+
  xlab("State of Employment")+ ylab("Salary ($)")

```


Attempting to get data in one datafile to plot into bar graph
```{r}

fm = distinct(ff, WORK_STATE, .keep_all = TRUE)
dfg = distinct(fg, WORK_STATE, .keep_all = TRUE)
total_dfm <- merge(fm,dfg,by="WORK_STATE")
total_dfms<-  subset(total_dfm, WORK_STATE!="Pennsylvania" & WORK_STATE!="North Carolina" & WORK_STATE!="Massachusetts")
```


Sum data analyst and data scientist wage for each state
```{r}
total_dfms$SUM <- rowSums(total_dfms[ ,c(3,5)])

#trying to plot this
ggplot(data=total_dfm, aes(x=WORK_STATE, y=SUM)) +
  geom_bar(stat="identity")
```


Tried this on tableau and wasn't getting the same results so had to figure out why I was getting a different result. 
Lets have another go at this
```{r}
#subsetting
data_interest = data.frame(salary_data_states[c("WORK_STATE", "JOB_TITLE_SUBGROUP", "PAID_WAGE_PER_YEAR")])
#data_int = na.exclude(data_interest)
data_interest$PAID_WAGE_PER_YEAR <- format(round(data_interest$PAID_WAGE_PER_YEAR, 2), nsmall = 2)
data_interest$PAID_WAGE_PER_YEAR <- as.numeric(data_interest$PAID_WAGE_PER_YEAR)
#data_int$JOB <- with(data_int, ifelse(JOB_TITLE_SUBGROUP == "data analyst", 1, 0))
#newdata <- data_interest[ which(data_interest$JOB_TITLE_SUBGROUP== c('data analyst', 'data scientist')),]

#newdata$PAID_WAGE_PER_YEAR <- as.numeric(newdata$PAID_WAGE_PER_YEAR)
#group by each state
#ndf = distinct(newdata, WORK_STATE, .keep_all = TRUE)

#sub-setting high paid states
#ndff <- ndf[ndf$WORK_STATE %in% c("California", "New York", "Washington", "New Jersey", "Texas", "Massachusetts") ,]

#ggplot(data=ndff, aes(x=ndf$WORK_STATE, y=ndf$PREVAILING_WAGE_PER_YEAR, fill=ndf$JOB_TITLE_SUBGROUP)) + geom_bar(stat="identity")
```

---------------------------------------------------------------------------
Let's try again for something clearer
```{r}

# I need to make a new dataset with column that has data analyst and data scientist in same column and prevailing wage in same column

#group <- newdata%>%group_by(WORK_STATE, JOB_TITLE_SUBGROUP)%>%summarise_all(funs(mean))
sea<-data_interest%>%group_by(WORK_STATE, JOB_TITLE_SUBGROUP)%>%summarise_all(funs(mean))
newdata <- sea[sea$JOB_TITLE_SUBGROUP %in% c('data analyst', 'data scientist'), ]

# Removing states I would not want to live in :
grouped<-  subset(newdata, WORK_STATE!="Louisana" & WORK_STATE!="Mississippi" & WORK_STATE!="New Mexico")


#select states with highest pay in data related jobs
sorted_total <- total_dfms[order(-total_dfms$SUM),]
table_all<-sorted_total[1:5 , ]
write.csv(table_all,file="/Users/LilyRoseSlater/Documents/Rwork/311/table_all.csv")
```

Here we can see that in combined data related jobs salary averages the highest paid states are California, Washington, Connecticut, Minnesota, New York

Lets put these together in a graph to commit to my branch:
```{r}

groupedstates <- grouped[grouped$WORK_STATE %in% c("California",  "Washington", "Connecticut", "Minnesota", "New York") ,]

  
 ggplot(data = groupedstates, aes(x = forcats::fct_reorder(WORK_STATE, PAID_WAGE_PER_YEAR, .fun = mean), y = PAID_WAGE_PER_YEAR, fill=groupedstates$JOB_TITLE_SUBGROUP)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  labs(x="States with Highest Avg Pay", y="Avg Paid Wage Per Year") +
  ggtitle("Which States have the Highest Average Data-related Salaries?")+
  guides(fill=guide_legend(title="And how do the data-\nrelated jobs differ?"))+
  scale_fill_manual(values=c('#003366', '#6699CC'))

```
This graph will cover the first two parts of the question. 
(What states (of those I am willing to move to) have the highest paying data-related salaries? Differences between job sub-categories?)

Because I used na.exclude after selecting my columns it was not changing the number of rows in the dataframe
Next question: Which companies have the highest salaries for those sub-types?

```{r}
companies = data.frame(salary_data_states[c("EMPLOYER_NAME", "WORK_STATE", "JOB_TITLE_SUBGROUP", "PAID_WAGE_PER_YEAR")])
comp = na.exclude(companies)
library(tm)
comp$EMPLOYER_NAME <-removePunctuation(comp$EMPLOYER_NAME)
#had to remove punctuation because companies had the same name written differently with commas

#inuit vs intuit intuit inc

comp$EMPLOYER_NAME[comp$EMPLOYER_NAME == "INTUIT INC"] <- "INTUIT"

comp$EMPLOYER_NAME[comp$EMPLOYER_NAME == "NETFLIX INC"] <- "NETFLIX"


sade<-comp%>%group_by(WORK_STATE, JOB_TITLE_SUBGROUP, EMPLOYER_NAME)%>%summarise_all(funs(mean))
comp_sade <- sade[sade$JOB_TITLE_SUBGROUP %in% c('data analyst', 'data scientist'), ]


#subset data jobs and sort by prevailing wage
#comp_sal <- comp[ which(comp$JOB_TITLE_SUBGROUP== c('data analyst', 'data scientist')),]

#comp_sal_sort <- comp_sal[order(-comp_sal$PAID_WAGE_PER_YEAR),]
#distinct company and group salary by wage for scientist and analyst
com_da = comp_sade[comp_sade$JOB_TITLE_SUBGROUP== c('data analyst'),]
com_ds = comp_sade[comp_sade$JOB_TITLE_SUBGROUP== c('data scientist'),]

com_da_sort <- com_da[order(-com_da$PAID_WAGE_PER_YEAR),]
com_ds_sort <- com_ds[order(-com_ds$PAID_WAGE_PER_YEAR),]

comp_states1 <- com_da[com_da$WORK_STATE %in% c("California",  "Washington", "Connecticut", "Minnesota", "New York") ,]

comp_states2 <- com_ds[com_ds$WORK_STATE %in% c("California",  "Washington", "Connecticut", "Minnesota", "New York") ,]


highest_da <- com_da_sort %>%                  # Company with the highest salary for data analysis in each state
  arrange(desc(PAID_WAGE_PER_YEAR)) %>% 
  group_by(WORK_STATE) %>%
  slice(1)


highest_ds <- com_ds_sort %>%                  # Company with the highest salary for data science in each state
  arrange(desc(PAID_WAGE_PER_YEAR)) %>% 
  group_by(WORK_STATE) %>%
  slice(1)

highest_states_da <- highest_da[order(-highest_da$PAID_WAGE_PER_YEAR),]
highest_states_ds <- highest_ds[order(-highest_ds$PAID_WAGE_PER_YEAR),]

comp_states1 <- highest_states_da[highest_states_da$WORK_STATE %in% c("California",  "Washington", "Connecticut", "Minnesota", "New York") ,]

comp_states2 <- highest_states_ds[highest_states_ds$WORK_STATE %in% c("California",  "Washington", "Connecticut", "Minnesota", "New York") ,]
```

Putting these into graphs
```{r}
ggplot(data=highest_states_da, aes(x= EMPLOYER_NAME, y= PAID_WAGE_PER_YEAR)) +
  geom_bar(stat="identity") +
  labs(x="Companies with Highest Pay", y="Prevailing Wage Per Year") +
  ggtitle("What companies have the Highest Paying Salaries for data analyst jobs?")

ggplot(data=highest_states_ds, aes(x= EMPLOYER_NAME, y= PAID_WAGE_PER_YEAR)) +
  geom_bar(stat="identity") +
  labs(x="Companies with Highest Pay", y="Paid Wage Per Year") +
  ggtitle("What companies have the Highest Paying Salaries for data scientist jobs?")
```

Putting this info into graphs doesn't add anything,
so instead going to commit the tables that have the top 6 companies for each job


```{r}

# head of each dataframe aka top 6 for each
kable(highest_states_da, caption = "Companies that pay the highest for Data Analyst positions in the 6 Highest Paying States")
kable(highest_states_ds, caption = "Companies that pay the highest for Data Scientist positions in the 6 Highest Paying States")
```


```{r}
 write.csv(highest_states_da,file="/Users/LilyRoseSlater/Documents/Rwork/311/highest_states_da.csv")

 write.csv(highest_states_ds,file="/Users/LilyRoseSlater/Documents/Rwork/311/highest_states_ds.csv")
```




## Extra Question

In the state with the highest salaries, which cities have the highest salaries? 
Are they in the same postal code? 

California is the highest paid for data analyst and scientists


```{r}
#sub-setting columns of interest
data_interest_cali = data.frame(salary_data_states[c("WORK_STATE", "WORK_CITY", "JOB_TITLE_SUBGROUP", "PAID_WAGE_PER_YEAR")])

#excluding rows with na's
data_int_cali = na.exclude(data_interest_cali)
str(data_int_cali)

cali_int = data_int_cali[data_int_cali$WORK_STATE == c("California" ), ]

cali_da = cali_int[cali_int$JOB_TITLE_SUBGROUP == c("data analyst" ), ]
cali_ds = cali_int[cali_int$JOB_TITLE_SUBGROUP == c( "data scientist"), ]

cali <- cali_int[ which(cali_int$JOB_TITLE_SUBGROUP== c('data analyst', 'data scientist')),]
#capitalizing all cities so they will group together
cali$WORK_CITY <- toupper(cali$WORK_CITY)
#Group together by city and sub-group
caligroup <- cali%>%group_by(WORK_CITY, JOB_TITLE_SUBGROUP)%>%summarise_all(funs(mean))

highest_cities_cali <-caligroup[order(-caligroup$PAID_WAGE_PER_YEAR),]
hccc <- head(highest_cities_cali)
cali_high_da = highest_cities_cali[highest_cities_cali$JOB_TITLE_SUBGROUP == c("data analyst" ), ]
cali_high_ds = highest_cities_cali[highest_cities_cali$JOB_TITLE_SUBGROUP == c("data scientist" ), ]
caligraph1da <- cali_high_da[1:5,]
caligraph1ds <- cali_high_ds[1:5,]

hcc <- highest_cities_cali[1:16,]
```

Graph this information

```{r}

ggplot(data=caligraph1da, aes(x= WORK_CITY, y= PAID_WAGE_PER_YEAR)) +
  geom_bar(stat="identity") +
  labs(x="Cities with Highest Pay", y="Prevailing Wage Per Year") +
  ggtitle("What cities have the Highest Paying Salaries for data analyst jobs?")

ggplot(data=caligraph1ds, aes(x= WORK_CITY, y= PAID_WAGE_PER_YEAR)) +
  geom_bar(stat="identity") +
  labs(x="Cities with Highest Pay", y="Prevailing Wage Per Year") +
  ggtitle("What cities have the Highest Paying Salaries for data scientist jobs?")


 ggplot(data = hcc, aes(x = forcats::fct_reorder(WORK_CITY, PAID_WAGE_PER_YEAR, .fun = mean), y = PAID_WAGE_PER_YEAR, fill=groupedstates$JOB_TITLE_SUBGROUP)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  labs(x="States with Highest Pay", y="Paid Wage Per Year") +
  ggtitle("Which States have the Highest Paying Data-related Salaries?")+
  guides(fill=guide_legend(title="And how do the data-\nrelated jobs differ?"))+
  scale_fill_manual(values=c('#003366', '#6699CC'))
```




